Assignment 2

### 2020-04-26 stuctbio/assignment2.tex
StuctBio: Assignment2: Included exercise question f and related python code, plus text improvements

% Reply log: RMSD between 20 pruned atom pairs is 1.037 angstroms; (across all 234 pairs: 22.983)
The c-RMSD as calculated by Chimera after superposition is $1.037$ \AA

\section*{b}

How to calculate the c-RMSD between two structures of different length? Homology? Alignment?

@@

We take the coodinates for each residue's backbone atoms from the \verb|.pdb| file and we calculate the Caley-Menger (border) matrix $B$. We get:

@@

By running the Python code \verb|numpy.linalg.matrix_rank(B)| we verify that the $rank(B) = 5$

Now in order to calculate the Gram matrix $G$, use the formula
\[
  G[i,j] := \frac{d_{i0}^2 - d_{ij}^2 + d_{j0}^2}{2}
\]
and while we can use the $B$ matrix for $\frac{d_{ij}^2}{2}$, in order to calculate $d_{i0}, d_{j0}$ we'll need to use the original coordinates. We therefore get
\begin{footnotesize}
  \[G =\]
  \[
    \begin{pmatrix}
      5199.442025 & 5188.332704 & 5287.185963 & 5326.089175 & 5419.655414 & 5410.547521 & 5409.324072 & 5401.249837 & 5487.240252\\
      5188.332704 & 5179.378713 & 5278.98134 & 5318.582438 & 5412.941379 & 5405.912056 & 5404.804417 & 5398.601585 & 5484.797241\\
      5287.185963 & 5278.98134 & 5380.931697 & 5421.417825 & 5518.00339 & 5511.837671 & 5511.00822 & 5505.626303 & 5593.808044\\
      5326.089175 & 5318.582438 & 5421.417825 & 5463.685574 & 5561.332566 & 5555.2185 & 5552.961168 & 5547.3885 & 5635.330409\\
      5419.655414 & 5412.941379 & 5518.00339 & 5561.332566 & 5661.125561 & 5655.807696 & 5653.583903 & 5648.730935 & 5738.403409\\
      5410.547521 & 5405.912056 & 5511.837671 & 5555.2185 & 5655.807696 & 5652.829306 & 5651.486458 & 5648.865404 & 5739.303511\\
      5409.324072 & 5404.804417 & 5511.00822 & 5552.961168 & 5653.583903 & 5651.486458 & 5651.920269 & 5650.381255 & 5742.063407\\
      5401.249837 & 5398.601585 & 5505.626303 & 5547.3885 & 5648.730935 & 5648.865404 & 5650.381255 & 5651.011947 & 5743.588506\\
      5487.240252 & 5484.797241 & 5593.808044 & 5635.330409 & 5738.403409 & 5739.303511 & 5742.063407 & 5743.588506 & 5838.531794
    \end{pmatrix}
  \]
\end{footnotesize}

and by using SVD we get the new coordinates
\begin{center}
  \[
    \begin{pmatrix}
      -0.720239668 & -3.29505773 & -1.06434145\\
      -0.719397583 & -1.95848694 & -0.462828815\\
      -0.733414994 & -1.34469699 & -0.384638014\\
      -0.739000952 & -1.33290190 & 0.827571807\\
      -0.752286576 & -0.772197814 & 1.08551735\\
      -0.751728771 & 0.755384109 & 1.13896831\\
      -0.751667006 & 1.37317774 & -0.0421124995\\
      -0.751193862 & 2.83776829 & -0.191918550\\
      -0.763282291 & 3.40972451 & -0.952375000
    \end{pmatrix}
  \]
\end{center}

@@

\textit{Note: Intuitively it seems it would make more sense to calculate $G$ using only the values from $B$. However when taking $d_{i0} = 1$ the c-RMSD becomes $\sim 5$.}

@@

We implement the perturbation by increasing all values in $B$ by a percentage.
Then we recalculate $G$, run SVD to get the $U$ and $\Sigma$ values, take the $S$ to be the $(3 \times 3)$ diagonal array with the three greatest values from $\Sigma$, and check the c-RMSD between the original cooredinates and the points from $\sqrt{S}U^T$.

The algorithm tried multiple points between $100\%$ and $200\%$ and then run again in the space where the c-RMSD would go over $1$. The conclusion was the the max percentage of perturbation for c-RMSD to be $< 1$ is $88.61\%$.

@@

Here for calculating the c-RMSD we will take only the distances that are less than a number $T$.
For the rest of the points we will assume that the distance is $0$ in order for it not to affect the final result.
We will also keep $R = 1$ \AA.

>>>>

% Reply log: RMSD between 20 pruned atom pairs is 1.037 angstroms; (across all 234 pairs: 22.983)
The c-RMSD as calculated by Chimera after superposition is
\begin{itemize}
  \item $1.037$ \AA \; between 20 pruned atom pairs
  \item $22.983$ \AA \; across all 234 pairs
\end{itemize}

%\section*{b}
%How to calculate the c-RMSD between two structures of different length? Homology? Alignment?

@@

We will take the coodinates for each residue's backbone atoms from the \verb|6FAT.pdb| file and we'll calculate the Caley-Menger (border) matrix $B$. We get:

@@

By running the Python code for $B$ \verb|numpy.linalg.matrix_rank(B)| we verify that $rank(B) = 5$.\\
Now in order to calculate the Gram matrix $G$, we will use the formula
\[
  G[i,j] := \frac{d_{i0}^2 - d_{ij}^2 + d_{j0}^2}{2}
\]
and while we can use the $B$ matrix to get $\frac{d_{ij}^2}{2}$, in order to calculate $d_{i0}$ and $d_{j0}$ we'll need to use the original coordinates. We therefore get:
\begin{footnotesize}
  \[G =\]
  \[
    \begin{pmatrix}
      5199.442025 & 5188.332704 & 5287.185963 & 5326.089175 & 5419.655414 & 5410.547521 & 5409.324072 & 5401.249837 & 5487.240252\\
      5188.332704 & 5179.378713 & 5278.98134 & 5318.582438 & 5412.941379 & 5405.912056 & 5404.804417 & 5398.601585 & 5484.797241\\
      5287.185963 & 5278.98134 & 5380.931697 & 5421.417825 & 5518.00339 & 5511.837671 & 5511.00822 & 5505.626303 & 5593.808044\\
      5326.089175 & 5318.582438 & 5421.417825 & 5463.685574 & 5561.332566 & 5555.2185 & 5552.961168 & 5547.3885 & 5635.330409\\
      5419.655414 & 5412.941379 & 5518.00339 & 5561.332566 & 5661.125561 & 5655.807696 & 5653.583903 & 5648.730935 & 5738.403409\\
      5410.547521 & 5405.912056 & 5511.837671 & 5555.2185 & 5655.807696 & 5652.829306 & 5651.486458 & 5648.865404 & 5739.303511\\
      5409.324072 & 5404.804417 & 5511.00822 & 5552.961168 & 5653.583903 & 5651.486458 & 5651.920269 & 5650.381255 & 5742.063407\\
      5401.249837 & 5398.601585 & 5505.626303 & 5547.3885 & 5648.730935 & 5648.865404 & 5650.381255 & 5651.011947 & 5743.588506\\
      5487.240252 & 5484.797241 & 5593.808044 & 5635.330409 & 5738.403409 & 5739.303511 & 5742.063407 & 5743.588506 & 5838.531794
    \end{pmatrix}
  \]
\end{footnotesize}

and by using SVD we get the new coordinates:
\begin{center}
  \[
    \begin{pmatrix}
      -0.720239668 & -3.29505773 & -1.06434145\\
      -0.719397583 & -1.95848694 & -0.462828815\\
      -0.733414994 & -1.34469699 & -0.384638014\\
      -0.739000952 & -1.33290190 & 0.827571807\\
      -0.752286576 & -0.772197814 & 1.08551735\\
      -0.751728771 & 0.755384109 & 1.13896831\\
      -0.751667006 & 1.37317774 & -0.0421124995\\
      -0.751193862 & 2.83776829 & -0.191918550\\
      -0.763282291 & 3.40972451 & -0.952375000
    \end{pmatrix}
  \]
\end{center}

@@

%\textit{Note: Intuitively it seems it would make more sense to calculate $G$ using only the values from $B$. However when taking $d_{i0} = 1$ the c-RMSD becomes $\sim 5$.}

@@

We implement the perturbation by increasing all values in $B$ by a percentage.
Then we recalculate $G$, run SVD to get the $U$ and $\Sigma$ values, take $S$ to be the $(3 \times 3)$ diagonal array with the three greatest values from $\Sigma$, and then check the c-RMSD between the original cooredinates and the points from $\sqrt{S}U^T$.

The algorithm tried 100 points between $1\%$ and $100\%$ (with a step of $1\%$) which was the incease applied to all values of $B$, and then run again for 10 more points in the space where the c-RMSD would go over $1$ to increase accuracy.

The conclusion was the max percentage of perturbation for c-RMSD to be less than $R = 1$ is around $88.61\%$.

@@

Here for calculating the c-RMSD we will take only the distances that are less than a number $T$.
For the rest of the points we will assume that the distance is $0$.
We will also keep $R = 1$ \AA.

\begin{center}
  \begin{tabular}{ ||c|c|| }
    \hline
    $T$ & c-RMSD\\
    \hline \hline
    30 & 4.848773458134514\\
    20 & 1.2941487875701954\\
    15 & 5.004258253300996\\
    10 & 5.243758773056963\\
    5 & 1.360263291246221\\
    4 & 1.3621607241108677\\
    3 & 1.3643224342718143\\
    2 & 1.15262288363306\\
    1 & 0.2292140768425968\\
    \hline
  \end{tabular}
\end{center}

And therefore for $T = 1$ we get the c-RMSD to be less than $R = 1$ \AA.
